Dynamics of Global and Local Social Influence

### 2021-09-01 manuscript_global_local_influence.Rmd
Added footnote & abstract

---
title: "Dynamics of Global and Local Social Influence"
author: "Fatima Alqabandi & Nicolas Restrepo"
date: '`r format(Sys.time(), "%d %B, %Y")`'
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
---

@@

```{r}
library(tidyverse)
library(igraph)
library(brms)
library(patchwork)
theme_set(theme_minimal())
```

@@

This paper has the following structure. First, we provide an overview of the theoretical issues that motivate this analysis. Second, we use the longitudinal network data-set of adolescent behavior collected by Knecht (2012) to suggest that measures of global and local influence might be differentially informative across different behaviors. Third, we build an agent-based model where actors use different combinations of strategies to learn from one another. The results show that different social learning strategies lead to varying diffusion trajectories. We close by discussing the implications of these findings for the study of cultural transmission.

@@

Peer effects are often explained by the need to “fit in” (Robalino and Macy 2018). Individuals might adopt new habits or beliefs not only because they are exposed to novel ideas, but also because certain lines of action fit better in the social contexts where they find themselves. Imagine Jane’s school friends start taking an interest in heavy metal and going to gigs. It makes sense for her to start going to those concerts as well. It is a behavior that fits her social context: it allows her to keep spending time with her friends. This view of social influence as a “problem of fit” (Fligstein 2001) underlies some of our most sophisticated attempts to disentangle how behaviors and ideas spread through networks. Whether it be actor-centric models of behavior dynamics (Snijders 2017) or spin-glass models of belief spread (Galesic and Stein 2019), the core assumption is the same: actors have a higher chance of adopting habits or beliefs if such change would bring about greater synchronicity between them and the people around them. Social influence then could be conceptualized as a problem of coordination (Fligstein 2001). We try to orient ourselves towards the lines of action we perceive as appropriate in our social contexts.

@@

This need to reconcile global and local forces is a central component of our widely accepted theories of social action (Abbott 2005; Martin 2003; Bourdieu 1990). Specific social settings reward certain behaviors and/or identities while discouraging others (Bourdieu 1990). For example, while a relentless drive to make profits might be celebrated in the corporate world, it would seem distasteful and unacceptable if embodied by an avant-garde artist. By virtue of inhabiting these social spaces, we align ourselves to their expectations and regularities. Through our piece-meal interactions, we build a picture of the patterns of behavior that configure the settings we navigate. We infer global patterns by way of our necessarily local experiences.

In the study of cultural diffusion, the interplay between local and global patterns is exemplified by ideas of “credibility” and “legitimacy” (Centola and Macy 2007; Rossman 2014). The former refers to the fact that agents might abstain from adopting an innovation until they see enough of their peers embracing it. The latter, in turn, reflects that behaviors that break expectations – e.g. a radical innovation or the disadoption of a widely accepted behavior – will be met with resistance in cultural transmission. Though implicitly, these concepts highlight the fact that, when deciding whether to adopt a behavior, agents take into account what their peers are doing but also try to infer more general patterns in the population. Agents aim to infer what the information to which they have access tells them about the general acceptance of a cultural trend in the overall population. To assess the credibility or the legitimacy of an innovation is partly about figuring out how the local patterns we observe fit into the wider global dynamics.

@@

Contemporary conceptions of social influence also wrestle with the tension between global and local dynamics. Increasingly, researchers have moved away from models that only consider immediate ties to more complex conceptualizations that specify varying sources of influence (DiMaggio and Garip 2012). For the purposes of this analysis, it is important to highlight the work that has shown that central nodes in a structure exercise considerable social influence
76 across the network (Robalino and Macy 2018; Paluck and Shepherd 2012). These findings lend credence to the argument that popular and salient individuals are particularly influential; a view that has been furthered in scholarly work (Sherif and Sherif 1964) but that has also been the backbone of policy interventions (Campbell et al. 2008). Peer effects then are not monolithic, but rather have varying weights depending on where – and who – they come from.

The underlying assumption here is that agents are capable of inferring who these central nodes are and of drawing information from them. This is not an unwarranted hypothesis.Humans are sophisticated social learners and one of the most widely used social learning strategies, across cultures, is to copy successful individuals (Jiménez and Mesoudi 2019). This means that we are highly sensitive to prestige cues. Experimental work has shown that even children engage in prestige-based social learning (Chudek et al. 2012). Salient individuals in a structure, then, are visible. They are visible because they have many connections – their in-degree count is high – but also because individuals are adept at inferring who has
high status in a given social group. As Shepherd (2017) notes, these processes of inference vary depending on the social positions that individuals occupy; nodes that are highly visible to a portion of the structure might be relatively obscure to others. But we still look for prestige cues, however limited our vision might be. Agents are not passive receptacles of social influence, but they instead follow cues that might indicate where to look for useful information.

There is an established line of interdisciplinary research that examines the different strategies that individuals use when foraging for information. Imitating “successful” individuals in the network proves an important strategy when tasks are complex or when environments are
highly variable (Henrich and Broesch 2011; McElreath et al. 2008; Henrich 2001). When
environments are relatively stable - e.g. Jane knows most people in her school will not show up tomorrow wearing Motorhead t-shirts - just looking at what the majority of group members are doing is an effective strategy (Richerson and Boyd 2008; Rendell et al. 2010). There are trade-offs between the different strategies: while identifying successful individuals entails knowledge about the markers of prestige in a given context, frequency-dependent learning, although efficient, can be ineffective in highly variable environments. Computational models and experimental data show that actors use these different strategies depending on the task at hand (McElreath et al. 2008; Barkoczi and Galesic 2016). Often the most successful strategies involve a blend of different heuristics, combined with some individual exploration.

@@

This line of inquiry is particularly promising because we have grounds to believe that different social learning strategies might shape how practices spread through a network. Edelmann (2018), for instance, shows that agents respond differently to peer influence depending on how culturally proximate they consider their alters and how culturally appropriate they deem the behavior in question. In other words, if a distant acquaintance of Jane recommends a country singer, she might just dismiss it. If one of her ‘metalhead’ friends provides the same recommendation, it might carry more weight just on account of how unexpected that would be. Computational work shows that these learning biases - such as preferentially attending to in-group members (Smaldino et al. 2017) or discarding information that is too distant from one’s existent preferences (Deffuant et al. 2000) - leads to varying diffusion trajectories. Even slight changes in the information that agents take into account give way to contrasting outcomes, from total consensus to fractured polarization. The question, then, is whether foraging for information from different sources, even when no biases are present, would also lead to varying patterns of diffusion. To tackle this issue, we need two things: first, statistical methods that can help us disentangle the importance of different sources of information for a given social phenomenon; second, ways of formalizing how agents use and give weight to different social learning strategies.

@@

To measure the differential impact of different sources of influence, we can draw ideas from research on social learning, specifically on the cultural evolution of strategies in games. When a soccer manager is deciding what formation to play, they need to consider many factors: their own track-record with a strategy, how that formation has fared in the league as a whole, and the kind of strategy their rivals are likely to use (Mesoudi 2020). Recent work explores how these global and local sources of information help us predict the popularity and evolution of certain strategies across time (Mesoudi 2020; Beheim, Thigpen, and McElreath 2014). Researchers have also designed games to conduct experimental work on broader social learning patterns (Deffner, Kleinow, and McElreath 2020). Though their statistical models vary in complexity, these studies tend to use repeated measures at the individual level to fit multilevel models that examine how informative personal record and social cues are when predicting what players do next. Our argument is that we can adapt the techniques that
these studies use in order to infer the informativeness of different sources of social influence in longitudinal network data.

In order to provide a formal model that specifies the importance of varying sources of influence, we can use as a starting point Friedkin and Johnsen’s (1997) classic formulation of social influence. Formalization in this case is useful because it specifies exactly how new ideas of social influence expand upon previous work, and - importantly - the assumptions we have to make in the process. Friedkin and Johnsen’s (1997) is as follows:

@@

In the Friedkin 170 and Johnsen (1997) model, the rows of the matrices must sum to one. Here, we have two popular nodes so they split the weight their actions or beliefs have. We can think about this as the visibility matrix: it simply stores information about the nodes whose opinions or behaviors agents consider when using prestige-biased social learning.

@@

$$ Y^{(t)} = \alpha Y^{(t-1)}[\lambda (W) + (1-\lambda)V] + (1-\alpha)Y^{(1)}$$
Though this equation might seem considerably longer than the original, the changes are
minimal. The term $\lambda$ tells us whether agents give more importance to the views of their peers or to those of popular nodes, and the matrix $V$ signals who the latter are. We can think of this specification as a more general form of the Friedkin and Johnsen (1997) model. If we set $\lambda$ to 1, then we get their seminal equation. The result then is a model of social influence that takes into account that individuals look at different places of the social structure when learning from one another.

@@

What we need is an empirical strategy to adjudicate the importance of different sources of social influence for a given social phenomenon. Consider a behavior of interest like drinking alcohol. The question we would like to answer is whether to predict agent j’s drinking habits at time t + 1, we should look at the behavior of their immediate ties or at the actions of the most popular nodes of the network at time t. Using longitudinal network data, we can generate predictions using only within-person variation in global and local influence (Beheim, Thigpen, and McElreath 2014). Thus, this approach is useful to disentangle how informative local and global dynamics are for explaining the spread and development of a phenomena of interest – like drinking – in networks across time. In the following analysis, we fit Bayesian multilevel models to longitudinal network data to show how the informativeness of different sources of influence varies across social phenomena.

Here, we use the longitudinal data-set of adolescent behaviors and opinions collected by Knecht et al. (2012). The data-set was collected over four waves in the space of three months between 2003 and 2004. In total, 2996 students – belonging to 14 schools in the Netherlands – were surveyed. 1509 students participated in all four waves and we focus on their responses for the following analysis.

@@

The explanatory variables are two measures we construct to represent the local and global prevalence of the behavior in question. The local score is the average behavior among ego’s friends. Thus, if ego has five nominated friends and two of them drink, then ego’s local score at that wave would be $\frac{2}{5}$. Local score, then, varies across respondents, reflecting differences in their immediate friendship networks. The global score, in turn, is the same for all individuals embedded in the same structure at a given wave. We calculate it by looking at the average behavior of the most popular nodes at a given time. To operationalize this idea of popularity, we look at eigen-vector centrality. We consider the nodes to be among the most popular if they have an eigen-vector centrality score above the 85th percentile[^There are different ways of operationalizing popularity; cf. (Robalino and Macy 2018). Here, we decide
to use a threshold that indexes a group of high-status, popular nodes. This captures how social learning scholars have discussed and formalized prestige-biased social learning (Jiménez and Mesoudi 2019). We want to reflect the idea that, when using this social learning strategy, agents look at a few individuals who theyconsider to be successful. Therefore, having a threshold that helps us distinguish between this group and therest of the nodes is appropriate.]. Say then that, in a given school, there are six nodes whose eigen-vector centrality is above that threshold, and one of them drinks. The global score for all nodes in that network in that wave would be $\frac{1}{6}$. Though these constructed variables are necessarily rough approximations, we consider that these scores can help us disentangle the importance of global and local sources of influence for different behaviors.

@@

Any examination about peer effects has to deal with the potential confounds of selection and influence (Steglich, Snijders, and Pearson 2010). The central danger is to think that individuals are becoming like their peers when really they are just choosing to interact with like-minded individuals in the first place. Here, longitudinal data can help us address that concern. This type of data provides repeated measures for each individual, which allows us to cluster estimates at the level of the respondent. This means we can use within-person variation to generate our predictions, each individual acting effectively as their own control group. The models above let us adjudicate whether changes in local and global scores, at the individual level, lead to differences in the likelihood of adopting a behavior of interest. Nonetheless, our main goal here is not to provide a rigorous causal account; Robalino and Macy (2018) have accomplished that well. Instead, we want to highlight two ideas: first, that we can disentangle the effects of central nodes and local ties; second, that the informativeness of these measures might vary for different social phenomena.

@@

We need to be careful when interpreting these coefficients. The coefficients of logit models are notoriously difficult to understand when examined separately. This is because the absolute probability of the outcome depends on all the moving parts of the model, not just the marginal distribution of a single coefficient estimate. Moreover, differences that might seem considerable between high values in the log-odds scale are smaller when transposed to the probability scale. For this reason, we plot the predicted probabilities of engaging in these behaviors, varying the values of global and local scores.

```{r, fig.height = 8, fig.width = 7}
bike_pp <- read_rds("Figures/bike_pp.R")
fare_pp <- read_rds("Figures/fare_pp.R")
copy_pp <- read_rds("Figures/copy_pp.R")
alco_pp <- read_rds("Figures/alcohol_pp.R")

patch <- (bike_pp / fare_pp / copy_pp / alco_pp)

patch
```

Figure 2 shows two density distributions for each model. We take the intercept of a participant in the sample and predict their probability of engaging in the behavior at hand at different values of the main independent variables. The red density plots, for instance, represent the predicted probability of engaging in the behavior when we add a one standard deviation increase to the global score while keeping local score at its median value. The blue density plots show the opposite: an increase in local score while holding global score constant. We notice that, in the model that predicts alcohol consumption, an increase in local score results in a higher probability of engaging in the behavior than a comparable increase in global score.The same occurs for the model that predicts illegaly copying a CD. For the remaining two models, the pattern is reversed: the distribution of predicted probabilities is higher after a one standard deviation increase in the global score. This means that a rise in global score is more likely to lead to higher predicted probability of engaging in the behavior.

Overall, the stories that the models tell are relatively consistent: the habits of friends and295of popular nodes are informative for understanding an agent’s probability of adopting the behavior. Increases in both global and local scores lead to a heightened predicted probability of engaging in all three behaviors. To an extent, this consistency is reassuring. The behaviors298at hand are all similar insofar as they are ‘deviant’ behaviors that require some risk-taking. Overwhelmingly different results would be concerning. Nonetheless, we do notice slight but considerable differences in how important each source of influence is for predicting the agents’ behaviors in the next wave.

@@

In this study, we emphasize that individuals are sophisticated social learners and that, for that reason, influence can come from different sources. This is not a novel idea; it is prevalent in the literature about social influence (Paluck and Shepherd 2012), cultural evolution (McElreath et al. 2008), and has even been the backbone of policy interventions Campbell394et al. 2008). Agents do not passively absorb the influence of their immediate ties but actively use learning heuristics, like following prestige-cues, to copy the successful. Our models of cultural transmission, nonetheless, do not tend to incorporate these varying learning heuristics or to explore how the weight we place on these sources of information affects how behaviors and ideas spread across networks. We contend that there are two main ways in which we can begin to tackle this issue: first, we can incorporate statistical techniques that are capable of adjudicating the importance of different sources of information in longitudinal network data; second, we can outline formal models that reflect that individuals look at different parts of the social structure when foraging for information.

@@

aking into account the weights placed on different sources of information is key because they shape trajectories of cultural transmission. The simulation results show that the importance that is given to social and individual information, as well as to local and global sources, shapes how and to what extent a behavior spreads across a network. When individuals pay little attention to social cues, there is predictably little transmission. As the weight of social cues increases, so does the reach of the behavior being transmitted. However, even when agents prioritize social cues, the places where they get these cues affect the processes of diffusion: a tendency to learn from popular nodes, for instance, tends to lead to quicker processes of transmission. How a behavior spreads across a network depends then, in part, on the learning techniques agents use to identify and imitate such action.

>>>>

---
title: "Dynamics of Global and Local Social Influence"
author: "Fatima Alqabandi & Nicolas Restrepo"
date: '`r format(Sys.time(), "%d %B, %Y")`'
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
bibliography: gli.bib
abstract: "When learning from one another, individuals use different strategies and heuristics. In this paper we argue that the social learning strategies agents use are important for understanding processes of cultural transmission. Even though research on social learning and social networks has increasingly considered different sources of social influence, many models of cultural transmission still tend to assume agents that learn from their peers in a single way. Here, we incorporate current knowledge about social learning into models of cultural transmission and assess its implications. We argue that classic formulations of social influence can be extended to incorporate different social learning heuristics. Then, using longitudinal network data, we show that there is evidence that agents place varying weight on different sources of social information, depending on the phenomenon at hand. We then build an agent-based simulation that incorporates the idea that individuals use different social learning strategies when deciding their lines of action. Our model is capable of replicating interesting findings, including the idea that sub-populations can emerge even in fully connected structures. The key is that we can arrive at that insight only by manipulating the weight that agents place on the different sources of social influence."
---

@@

```{r}
library(tidyverse)
library(igraph)
library(brms)
library(patchwork)
library(panelr)
theme_set(theme_classic())
```

@@

This paper has the following structure. First, we provide an overview of the theoretical issues that motivate this analysis. Second, we use the longitudinal network data-set of adolescent behavior collected by Knecht [-@knechtNetworkActorAttributes2012] to suggest that measures of global and local influence might be differentially informative across different behaviors. Third, we build an agent-based model where actors use different combinations of strategies to learn from one another. The results show that different social learning strategies lead to varying diffusion trajectories. We close by discussing the implications of these findings for the study of cultural transmission.

@@

Peer effects are often explained by the need to “fit in” [@robalinoPeerEffectsAdolescent2018]. Individuals might adopt new habits or beliefs not only because they are exposed to novel ideas, but also because certain lines of action fit better in the social contexts where they find themselves. Imagine Jane’s school friends start taking an interest in heavy metal and going to gigs. It makes sense for her to start going to those concerts as well. It is a behavior that fits her social context: it allows her to keep spending time with her friends. This view of social influence as a “problem of fit” [@fligsteinSocialSkillTheory2001] underlies some of our most sophisticated attempts to disentangle how behaviors and ideas spread through networks. Whether it be actor-centric models of behavior dynamics [@snijdersStochasticActororientedModels2017] or spin-glass models of belief spread [@galesicStatisticalPhysicsModels2019], the core assumption is the same: actors have a higher chance of adopting habits or beliefs if such change would bring about greater synchronicity between them and the people around them. Social influence then could be conceptualized as a problem of coordination [@fligsteinSocialSkillTheory2001]. We try to orient ourselves towards the lines of action we perceive as appropriate in our social contexts.

@@

This need to reconcile global and local forces is a central component of our widely accepted theories of social action [@abbottLinkedEcologiesStates2005; @martinWhatFieldTheory2003; @bourdieuLogicPractice1990]. Specific social settings reward certain behaviors and/or identities while discouraging others [@bourdieuLogicPractice1990]. For example, while a relentless drive to make profits might be celebrated in the corporate world, it would seem distasteful and unacceptable if embodied by an avant-garde artist. By virtue of inhabiting these social spaces, we align ourselves to their expectations and regularities. Through our piece-meal interactions, we build a picture of the patterns of behavior that configure the settings we navigate. We infer global patterns by way of our necessarily local experiences.

In the study of cultural diffusion, the interplay between local and global patterns is exemplified by ideas of “credibility” and “legitimacy” [@centolaComplexContagionsWeakness2007; @rossmanDiffusionLegitimateDiffusion2014]. The former refers to the fact that agents might abstain from adopting an innovation until they see enough of their peers embracing it. The latter, in turn, reflects that behaviors that break expectations – e.g. a radical innovation or the disadoption of a widely accepted behavior – will be met with resistance in cultural transmission. Though implicitly, these concepts highlight the fact that, when deciding whether to adopt a behavior, agents take into account what their peers are doing but also try to infer more general patterns in the population. Agents aim to infer what the information to which they have access tells them about the general acceptance of a cultural trend in the overall population. To assess the credibility or the legitimacy of an innovation is partly about figuring out how the local patterns we observe fit into the wider global dynamics.

@@

Contemporary conceptions of social influence also wrestle with the tension between global and local dynamics. Increasingly, researchers have moved away from models that only consider immediate ties to more complex conceptualizations that specify varying sources of influence [@dimaggioNetworkEffectsSocial2012]. For the purposes of this analysis, it is important to highlight the work that has shown that central nodes in a structure exercise considerable social influence
76 across the network [@robalinoPeerEffectsAdolescent2018; @paluckSalienceSocialReferents2012]. These findings lend credence to the argument that popular and salient individuals are particularly influential; a view that has been furthered in scholarly work [@sherifReferenceGroupsExploration1964] but that has also been the backbone of policy interventions [@campbellInformalSchoolbasedPeerled2008]. Peer effects then are not monolithic, but rather have varying weights depending on where – and who – they come from.

The underlying assumption here is that agents are capable of inferring who these central nodes are and of drawing information from them. This is not an unwarranted hypothesis.Humans are sophisticated social learners and one of the most widely used social learning strategies, across cultures, is to copy successful individuals [@jimenezPrestigebiasedSocialLearning2019]. This means that we are highly sensitive to prestige cues. Experimental work has shown that even children engage in prestige-based social learning [@chudekPrestigebiasedCulturalLearning2012]. Salient individuals in a structure, then, are visible. They are visible because they have many connections – their in-degree count is high – but also because individuals are adept at inferring who has high status in a given social group. As Shepherd [-@shepherdStructurePerceptionHow2017] notes, these processes of inference vary depending on the social positions that individuals occupy; nodes that are highly visible to a portion of the structure might be relatively obscure to others. But we still look for prestige cues, however limited our vision might be. Agents are not passive receptacles of social influence, but they instead follow cues that might indicate where to look for useful information.

There is an established line of interdisciplinary research that examines the different strategies that individuals use when foraging for information. Imitating “successful” individuals in the network proves an important strategy when tasks are complex or when environments are highly variable [@henrichNatureCulturalTransmission2011; @mcelreathExistenceAimingOutside2008; @henrichCulturalTransmissionDiffusion2001]. When environments are relatively stable - e.g. Jane knows most people in her school will not show up tomorrow wearing Motorhead t-shirts - just looking at what the majority of group members are doing is an effective strategy [@richersonNotGenesAlone2008; @rendellWhyCopyOthers2010]. There are trade-offs between the different strategies: while identifying successful individuals entails knowledge about the markers of prestige in a given context, frequency-dependent learning, although efficient, can be ineffective in highly variable environments. Computational models and experimental data show that actors use these different strategies depending on the task at hand [@mcelreathExistenceAimingOutside2008; @barkocziSocialLearningStrategies2016]. Often the most successful strategies involve a blend of different heuristics, combined with some individual exploration.

@@

This line of inquiry is particularly promising because we have grounds to believe that different social learning strategies might shape how practices spread through a network. Edelmann [-@edelmannFormalizingSymbolicBoundaries2018], for instance, shows that agents respond differently to peer influence depending on how culturally proximate they consider their alters and how culturally appropriate they deem the behavior in question. In other words, if a distant acquaintance of Jane recommends a country singer, she might just dismiss it. If one of her ‘metalhead’ friends provides the same recommendation, it might carry more weight just on account of how unexpected that would be. Computational work shows that these learning biases - such as preferentially attending to in-group members [@smaldinoAdoptionSocialMarker2017] or discarding information that is too distant from one’s existent preferences [@deffuantMixingBeliefsInteracting2000] - leads to varying diffusion trajectories. Even slight changes in the information that agents take into account give way to contrasting outcomes, from total consensus to fractured polarization. The question, then, is whether foraging for information from different sources, even when no biases are present, would also lead to varying patterns of diffusion. To tackle this issue, we need two things: first, statistical methods that can help us disentangle the importance of different sources of information for a given social phenomenon; second, ways of formalizing how agents use and give weight to different social learning strategies.

@@

To measure the differential impact of different sources of influence, we can draw ideas from research on social learning, specifically on the cultural evolution of strategies in games. When a soccer manager is deciding what formation to play, they need to consider many factors: their own track-record with a strategy, how that formation has fared in the league as a whole, and the kind of strategy their rivals are likely to use [@mesoudiCulturalEvolutionFootball2020[. Recent work explores how these global and local sources of information help us predict the popularity and evolution of certain strategies across time [@mesoudiCulturalEvolutionFootball2020; @beheimStrategicSocialLearning2014]. Researchers have also designed games to conduct experimental work on broader social learning patterns [@deffnerDynamicSocialLearning2020]. Though their statistical models vary in complexity, these studies tend to use repeated measures at the individual level to fit multilevel models that examine how informative personal record and social cues are when predicting what players do next. Our argument is that we can adapt the techniques that these studies use in order to infer the informativeness of different sources of social influence in longitudinal network data.

In order to provide a formal model that specifies the importance of varying sources of influence, we can use as a starting point Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] classic formulation of social influence. Formalization in this case is useful because it specifies exactly how new ideas of social influence expand upon previous work, and - importantly - the assumptions we have to make in the process. Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] is as follows:

@@

In the Friedkin and Johnsen [-@friedkinSocialPositionsInfluence1997] model, the rows of the matrices must sum to one. Here, we have two popular nodes so they split the weight their actions or beliefs have. We can think about this as the visibility matrix: it simply stores information about the nodes whose opinions or behaviors agents consider when using prestige-biased social learning.

@@

$$ Y^{(t)} = \alpha Y^{(t-1)}[\lambda (W) + (1-\lambda)V] + (1-\alpha)Y^{(1)}$$
Though this equation might seem considerably longer than the original, the changes are
minimal. The term $\lambda$ tells us whether agents give more importance to the views of their peers or to those of popular nodes, and the matrix $V$ signals who the latter are. We can think of this specification as a more general form of the Friedkin and Johnsen [-@friedkinSocialPositionsInfluence1997] model. If we set $\lambda$ to 1, then we get their seminal equation. The result then is a model of social influence that takes into account that individuals look at different places of the social structure when learning from one another.

@@

What we need is an empirical strategy to adjudicate the importance of different sources of social influence for a given social phenomenon. Consider a behavior of interest like drinking alcohol. The question we would like to answer is whether to predict agent j’s drinking habits at time t + 1, we should look at the behavior of their immediate ties or at the actions of the most popular nodes of the network at time t. Using longitudinal network data, we can generate predictions using only within-person variation in global and local influence [@beheimStrategicSocialLearning2014]. Thus, this approach is useful to disentangle how informative local and global dynamics are for explaining the spread and development of a phenomena of interest – like drinking – in networks across time. In the following analysis, we fit Bayesian multilevel models to longitudinal network data to show how the informativeness of different sources of influence varies across social phenomena.

Here, we use the longitudinal data-set of adolescent behaviors and opinions collected by Knecht et al. [-@knechtNetworkActorAttributes2012]. The data-set was collected over four waves in the space of three months between 2003 and 2004. In total, 2996 students – belonging to 14 schools in the Netherlands – were surveyed. 1509 students participated in all four waves and we focus on their responses for the following analysis.

@@

The explanatory variables are two measures we construct to represent the local and global prevalence of the behavior in question. The local score is the average behavior among ego’s friends. Thus, if ego has five nominated friends and two of them drink, then ego’s local score at that wave would be $\frac{2}{5}$. Local score, then, varies across respondents, reflecting differences in their immediate friendship networks. The global score, in turn, is the same for all individuals embedded in the same structure at a given wave. We calculate it by looking at the average behavior of the most popular nodes at a given time. To operationalize this idea of popularity, we look at eigen-vector centrality. We consider the nodes to be among the most popular if they have an eigen-vector centrality score above the 85th percentile[^There are different ways of operationalizing popularity; cf. [@robalinoPeerEffectsAdolescent2018]][^We conducted the same analyses using the 80th and 90th percentile as cutoffs and got largely consistent results]. Here, we decide
to use a threshold that indexes a group of high-status, popular nodes. This captures how social learning scholars have discussed and formalized prestige-biased social learning [@jimenezPrestigebiasedSocialLearning2019]. We want to reflect the idea that, when using this social learning strategy, agents look at a few individuals who they consider to be successful. Therefore, having a threshold that helps us distinguish between this group and the rest of the nodes is appropriate. Say then that, in a given school, there are six nodes whose eigen-vector centrality is above that threshold, and one of them drinks. The global score for all nodes in that network in that wave would be $\frac{1}{6}$. Though these constructed variables are necessarily rough approximations, we consider that these scores can help us disentangle the importance of global and local sources of influence for different behaviors.

@@

Any examination about peer effects has to deal with the potential confounds of selection and influence [@steglichDynamicNetworksBehavior2010]. The central danger is to think that individuals are becoming like their peers when really they are just choosing to interact with like-minded individuals in the first place. Here, longitudinal data can help us address that concern. This type of data provides repeated measures for each individual, which allows us to cluster estimates at the level of the respondent. This means we can use within-person variation to generate our predictions, each individual acting effectively as their own control group. The models above let us adjudicate whether changes in local and global scores, at the individual level, lead to differences in the likelihood of adopting a behavior of interest. Nonetheless, our main goal here is not to provide a rigorous causal account; Robalino and Macy [-@robalinoPeerEffectsAdolescent2018] have accomplished that well. Instead, we want to highlight two ideas: first, that we can disentangle the effects of central nodes and local ties; second, that the informativeness of these measures might vary for different social phenomena.

@@

We need to be careful when interpreting these coefficients. The coefficients of logit models are notoriously difficult to understand when examined separately. This is because the absolute probability of the outcome depends on all the moving parts of the model, not just the marginal distribution of a single coefficient estimate. Moreover, differences that might seem considerable between high values in the log-odds scale are smaller when transposed to the probability scale. For this reason, it is key to look at the predicted probabilities of engaging in the behaviors in question. Here, we present predicted probabilities for running a red light while biking and for drinking alcohol.

```{r}
theme_set(theme_classic())
set.seed(33)
#### Alcohol ####
alcohol_model <- readRDS('Analyses/alcohol_model.rds')
# Import the data 
df_wide <- read_csv('Data/alcohol_wide.csv')
# Turn into a long panel dataset 
df_long <- long_panel(df_wide, 
                      id = "id", 
                      prefix = "_w", 
                      begin = 1, 
                      end = 4, 
                      label_location = "end")

# Rename, change, and clean 
dfl <- df_long %>% 
  mutate_at(.vars = c("friends", 
                      "popular", 
                      "actalcoh"), 
            as.numeric) %>% 
  filter(!is.na(friends) & 
           !is.nan(friends) & 
           !is.na(popular) & 
           !is.nan(popular))

dfl_scaled <- dfl
dfl_scaled$friends_z <- (dfl_scaled$friends-mean(dfl_scaled$friends)/sd(dfl_scaled$friends))
dfl_scaled$popular_z <- (dfl_scaled$popular-mean(dfl_scaled$popular)/sd(dfl_scaled$popular))

dfl_scaled <- dfl_scaled %>% 
  panel_data(., id = id, wave = wave)

median_popular <- median(dfl_scaled$popular_z)
median_friend <- median(dfl_scaled$friends_z)
one_sd_popular <- median_popular+sd(dfl_scaled$popular)
one_sd_friend<- median_friend+sd(dfl_scaled$friends_z)

person <- sample(dfl_scaled$id, 50)
newdata <- data.frame(id = rep(person, 2), 
                      friends_z = rep(c(median_friend, one_sd_friend), each = 50), 
                      popular_z = rep(c(one_sd_popular, median_popular), each = 50), 
                      wave = rep(2, 100),
                      type = rep(c("global score", "local score"), each = 50))

preds <- predict(alcohol_model, 
                 newdata = newdata)

newdata$predicted_probs <- preds[,1]

alc <- newdata %>% 
  ggplot(aes(x = type, 
             y = predicted_probs, 
             group = id)) +
  geom_line(alpha = 0.7) + 
  labs(title = "Alcohol", 
       y = "Predicted probability", 
       x = "Type of increase")

#### Running red light ####

dfl_scaled <- read_csv('Data/bike_std.csv')
bike_model <- readRDS('Analyses/bike_model.rds')

median_popular <- median(dfl_scaled$popular_z)
median_friend <- median(dfl_scaled$friends_z)
one_sd_popular <- median_popular+sd(dfl_scaled$popular)
one_sd_friend<- median_friend+sd(dfl_scaled$friends_z)

person <- sample(dfl_scaled$id, 50)
newdata <- data.frame(id = rep(person, 2), 
                      friends_z = rep(c(median_friend, one_sd_friend), each = 50), 
                      popular_z = rep(c(one_sd_popular, median_popular), each = 50), 
                      wave = rep(2, 100),
                      type = rep(c("global score", "local score"), each = 50))

preds <- predict(bike_model, 
                 newdata = newdata)

newdata$predicted_probs <- preds[,1]

bike <- newdata %>% 
  ggplot(aes(x = type, 
             y = predicted_probs, 
             group = id)) +
  geom_line(alpha = 0.7) + 
  labs(title = "Running a red light", 
       y = "Predicted probability", 
       x = "Type of increase", 
       caption = "Figure 2")

(alc | bike)
```

Figure 2 shows two predicted probabilities for each model. We randomly sample 50 agents and predict their probability of engaging in the behavior at hand at different values of the main independent variables. The leftmost point of each plot represents the predicted probability of engaging in the behavior when we add a one standard deviation increase to the global score while keeping local score at its median value. The rightmost point shows the opposite: an increase in local score while holding global score constant. We notice that, in the model that predicts alcohol consumption, there is a clear concentration among lower probabilities; most respondents have a low probability of drinking. However, in many cases, an increase in local score results in a slightly higher probability of engaging in the behavior than a comparable increase in global score. For running a red right while biking, the pattern is reversed. There is a concentration at high probabilities, indicating that most agents have recently ran a red light. The predicted probabilities are slightly higher after a one standard deviation increase in the global score. This means that a rise in global score is more likely to lead to higher predicted probability of engaging in the behavior.

Overall, the stories that the models tell are relatively consistent: the habits of friends and of popular nodes are informative for understanding an agent’s probability of adopting the behavior. Increases in both global and local scores lead to a heightened predicted probability of engaging in all three behaviors. To an extent, this consistency is reassuring. The behaviors at hand are all similar insofar as they are ‘deviant’ behaviors that require some risk-taking. Overwhelmingly different results would be concerning. Nonetheless, we do notice slight but considerable differences in how important each source of influence is for predicting the agents’ behaviors in the next wave.

@@

In this study, we emphasize that individuals are sophisticated social learners and that, for that reason, influence can come from different sources. This is not a novel idea; it is prevalent in the literature about social influence [@paluckSalienceSocialReferents2012], cultural evolution [@mcelreathExistenceAimingOutside2008], and has even been the backbone of policy interventions [@campbellInformalSchoolbasedPeerled2008]. Agents do not passively absorb the influence of their immediate ties but actively use learning heuristics, like following prestige-cues, to copy the successful. Our models of cultural transmission, nonetheless, do not tend to incorporate these varying learning heuristics or to explore how the weight we place on these sources of information affects how behaviors and ideas spread across networks. We contend that there are two main ways in which we can begin to tackle this issue: first, we can incorporate statistical techniques that are capable of adjudicating the importance of different sources of information in longitudinal network data; second, we can outline formal models that reflect that individuals look at different parts of the social structure when foraging for information.

@@

Taking into account the weights placed on different sources of information is key because they shape trajectories of cultural transmission. The simulation results show that the importance that is given to social and individual information, as well as to local and global sources, shapes how and to what extent a behavior spreads across a network. When individuals pay little attention to social cues, there is predictably little transmission. As the weight of social cues increases, so does the reach of the behavior being transmitted. However, even when agents prioritize social cues, the places where they get these cues affect the processes of diffusion: a tendency to learn from popular nodes, for instance, tends to lead to quicker processes of transmission. How a behavior spreads across a network depends then, in part, on the learning techniques agents use to identify and imitate such action.
